Improving the Image Acquisition Rate \\ of an Atomic Force Microscope \\ Through Sub-sampling and Reconstruction

### 2019-06-28 latex/afm-cs-main.tex
Update abstract and conclusion.

\maketitle
\begin{abstract}
  Undersampling-based approaches in Atomic Force Microscopy (AFM) aim
  to reduce the time to acquire an image by reducing the number of
  measurements needed while still maintaining image quality. The approach consists of two components: data acquisition and image reconstruction. While the underlying idea is straightforward, practical implementation of a working scheme involves solving a variety of non-trivial problems. In this
  paper, we describe one such hardware implementation and demonstration based on the use of collections of short scans
  known as $\mu$--paths. Using a commercial AFM, we acquire data on a
  grating sample. Reconstructions are made from these data using
  % three approaches: (1) inpainting, an interpolation technique that
  % diffuses information from sampled locations into unsampled
  % regions, (2)
  % basis-pursuit, a compressive sensing-based algorithm that seeks to
  % minimize a
  % measure of sparsity in the underlying image, and (3)
  a new variant of basis pursuit, Basis Pursuit with Vertical
  Variation, that is designed to reduce artifacts arising from
  the sampling pattern. The quality of the resulting images is
  compared to images from standard raster scans of the same regions at
  comparable imaging rates using both the peak signal-to-noise ratio
  and the structured similarity index metric and a new metric we call
  the relative damage index. These images are compared both to standard raster scans as well as to simple subline sampling where only a subset of the rows of the image are acquired (followed by reconstruction). \textcolor{red}{These experiments demonstrate that at slow scan
  rates, the $\mu-$path scheme produces similar image quality with
  significantly less sampling (and thus less tip-sample interaction)
  and that at high rates, the undersampling scheme produces higher
  quality images than raster scanning.}
\end{abstract}

@@

Our previous work on the $\mu$-path pattern, using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are as follows.
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype \cite{braker_hardware_2018}, we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an effect which neither our prior experimental work nor simulation studies accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to coarse raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a coarse raster scan is that it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{luo2015comparison,maxwell2014compressed} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
\end{itemize}

@@

% \begin{figure*}
%   % \centering
%   \includesvg[scale=.98]{figures/cs_raster_images_6-5-2019.svg}
%   \caption{Raster and compressed sensing images of a 5 micron by 5 micron 
%     area of the CS-20NG grating. All images are 512$\times$512 pixels.
%     The rows of pixels indicated by the red lines are
%     shown in Fig.~\ref{fig:pixel_rows}.}
%   \label{fig:resultsF1_images}
% \end{figure*}

% \begin{table}[t!]
%   \centering
%   \caption{Performance metrics for raster and CS scans.}
%   \begin{tabular}{cccccc}
%     \input{tables/cs_raster_table_6-5-2019_muInf_dct2.tex}
%   \end{tabular}
%   \label{tab:rast_vs_cs_v1}
% \end{table}

@@

\begin{figure}
  \includesvg[width=1\columnwidth]{figures/cs_raster_pixel_rows_6-5-2019.svg}
  \caption{Rows of pixels, as indicated by the red lines in
    Fig.~\ref{fig:resultsF1_images}. For clarity, not all images are
    included.}
  \label{fig:pixel_rows}
\end{figure}
For each raster scan and each CS scan, we computed the SSIM and PSNR
using the 1~Hz raster scan as the master and also computed the RDI
metric. These are plotted against total acquisition time in
Fig.~\ref{fig:time_vs_metrics}. Unfortunately, the
SSIM and PSNR metrics are not particularly illuminating. There are
several inconsistencies. For example, it makes no sense that the 1~Hz CS
scan at 25\% has a \emph{worse} PSNR and SSIM than the 5~Hz scan; nor
does it make sense that the 8~Hz full raster scan has a worse SSIM and
PSNR than the 10~Hz full raster scan.

@@

\begin{figure}
  \includesvg[width=1\columnwidth]{figures/improvements.svg}
  \caption{Fraction of improvement in total scan time over a scan with
    512 lines vs scan rate for the three sets of CS scans (at 25\%,
    15\%, and 12.5\%) and coarser raster scans at 128 and 64 lines.}
  \label{fig:improve_512}
\end{figure}

\section{Conclusions}\label{sec:conclusions}
The main conclusion of this paper is that with CS, it seems that we
can acquire images faster yet with less specimen damage than is
possible while raster scanning a full image. On the other hand, it
appears that reconstructing a coarse raster scan makes more sense for
faster scan rates. This is notable, since a coarse raster scan only
requires an additional step in post-processing. It remains to be seen
if this conclusion holds for a wide range of instruments.

%%%%%%%%% then the Bibliography, if any %%%%%%%%%
% \bibliographystyle{plain} % or "siam", or "alpha", etc.
\bibliographystyle{IEEEtran} % or "siam", or "alpha", etc.
\bibliography{braker_bibs,Andersson_bibs,Yufans_bibs} % Bib database in "refs.bib"

\begin{IEEEbiography}
  [{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{figures/braker_headshot.jpg}}]
  {Roger A. Braker} earned the B.A. degree in physics from the University of Oklahoma in 2005 and the M.S. degree in electrical engineering from the University of Colorado in 2017.
  Since 2014, he has been a graduate student in the Electrical, Computer, and Energy Engineering
  Department at the University of Colorado Boulder.

\end{IEEEbiography}

\begin{IEEEbiography}
  [{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{figures/yufan.JPG}}]
  {Yufan Luo}
received dual B.S. degrees in automation and in management science from Zhejiang University, Zhejiang, China, in 2011, an M.S. degree in industrial and operations engineering from the University of Michigan, Ann Arbor, MI, USA, in 2012, and a Ph.D. degree in systems engineering from Boston University in 2019. He is currently with Juniper Networks, Sunnyvale, CA.

\end{IEEEbiography}

\begin{IEEEbiography}
  [{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{figures/pao_headshot.jpg}}]
  {Lucy Y. Pao} 
  is the Palmer Endowed Chair Professor in the Electrical, Computer, and Energy Engineering Department at the University of Colorado Boulder. She earned B.S., M.S., and Ph.D. degrees in Electrical Engineering from Stanford University. Her research has primarily focused on engineering control systems, with applications ranging from atomic force microscopes to multi-megawatt wind energy systems. She is a Fellow of the IEEE and the International Federation of Automatic Control (IFAC). Selected recent awards include the 2012 IEEE Control Systems Magazine Outstanding Paper Award, the 2015 Society for Industrial and Applied Mathematics (SIAM) Journal on Control and Optimization Best Paper Prize, the 2017 Control Engineering Practice Award from the American Automatic Control Council, and the Scientific Award 2017 from the European Academy of Wind Energy. Selected recent and current professional society activities include being a Fellow of the Renewable and Sustainable Energy Institute (2009-present), 
  General Chair of the 2013 American Control Conference, member of the IEEE Control Systems Society (CSS) Board of Governors (2011-2013 and 2015), IEEE CSS Fellow Nominations Chair (2016-2018), and member of the IFAC Executive Board (2017-2020).

\end{IEEEbiography}

\begin{IEEEbiography}
  [{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{figures/SeanAndersson_IEEE.jpg}}]
  {Sean B. Andersson}
received the B.S. degree in engineering and applied physics from Cornell University, Ithaca, NY, USA, in 1994, the M.S. degree in mechanical engineering from Stanford University, Stanford, CA, USA, in 1995, and the Ph.D. degree in electrical and computer engineering from the University of Maryland, College Park, MD, USA, in 2003. He has worked at AlliedSignal Aerospace and Aerovironment, Inc. and is currently an Associate Professor of mechanical engineering and of systems engineering with Boston University, Boston, MA, USA. His research interests include systems and control theory with applications in scanning probe microscopy, dynamics in molecular systems, and robotics. He received an NSF CAREER award in 2009, is a senior member of the IEEE, and was an associate editor for the IEEE Transactions on Automatic Control (2014-2018) and for the SIAM Journal on Control and Optimization (2013-2018).
\end{IEEEbiography}

>>>>

\maketitle
\begin{abstract}
  Undersampling-based approaches in Atomic Force Microscopy (AFM) aim
  to reduce the time to acquire an image by reducing the number of
  measurements needed while still maintaining image quality. The approach consists of two components: data acquisition and image reconstruction. While the underlying idea is straightforward, practical implementation of a working scheme involves solving a variety of non-trivial problems. In this
  paper, we describe one such hardware implementation and demonstration based on the use of collections of short scans
  known as $\mu$--paths. Using a commercial AFM, we acquire data on a
  grating sample. Reconstructions are made from these data using
  a new variant of basis pursuit, Basis Pursuit with Vertical
  Variation, that is designed to reduce artifacts arising from
  the sampling pattern. The quality of the resulting images is
  compared to images from standard raster scans of the same regions at
  comparable imaging rates using both the peak signal-to-noise ratio
  and the structured similarity index metric and a new metric we call
  the relative damage index. These images are compared (across a range of scan speeds)
  to both standard raster
  scans as well as to simple subline sampling where only a subset of the
  rows of the image are acquired (followed by reconstruction).
  These experiments show that at low sampling densities and slow scan rates,
  $\mu$-path scanning achieves better image quality than subline sampling.
  Conversely, for faster scan rates, subline sampling achieves better image quality
  in less time.
  % These experiments demonstrate that at slow scan
  % rates, the $\mu-$path scheme produces similar image quality with
  % significantly less sampling (and thus less tip-sample interaction)
  % and that at high rates, the undersampling scheme produces higher
  % quality images than raster scanning.
\end{abstract}

@@

Our previous work on the $\mu$-path pattern, using theoretical
calculations, simulations, and a preliminary implementation,
demonstrates that the approach can achieve scanning time
reduction while maintaining faithful image reconstruction
\cite{maxwell2014compressed,Luo:2015tu, braker_hardware_2018}. 
The main contributions of this paper relative to that past work are as follows.
% \begin{enumerate}[wide, labelwidth=!, labelindent=0pt]
\begin{itemize}
\item We introduce a new reconstruction algorithm which is designed to
minimize the artifacts arising from the structure of the $\mu$-path
scanning pattern. This algorithm is described and compared to existing
reconstruction methods through simulations in Sec. \ref{sec:reconstructionMethods}.
\item In our prior experimental prototype \cite{braker_hardware_2018}, we used very basic integral controllers for all three axes for both raster and $\mu$-path scans. Here, we design compensators for each axis using standard loop-shaping techniques (Sec.~\ref{sec:control}), which result in a substantial increase in closed-loop bandwidth (e.g., 50 Hz to 450~Hz for the $Z$-axis).
We show in Sec.~\ref{sec:ff_control} that one of the limitations of $\mu$-path scanning is the discontinuous $X$-axis reference (e.g., in moving between $\mu$-paths) which excites the cross-coupling modes between the $X$ and $Z$ axes, an effect which neither our prior experimental work nor simulation studies accounted for. We mitigate the cross coupling with a feed-forward design.
\item We experimentally compare (Sec. \ref{sec:results}) $\mu$-path scanning to not only full density raster scans but also to coarse raster scans which are interpolated to a full image. This is in contrast to typical sub-line sampling which takes scan lines at random. The clear advantage to a coarse raster scan is that it requires no modification to the instrument, only to the post-processing.
\item The bulk of this work is based on experiment, not simulation, in contrast to prior work by ourselves \cite{luo2015comparison,maxwell2014compressed} and others \cite{han_optimal_2018,jensen_reconstruction_2013,oxvig_structure_2017}. Our main focus here is on delineating the practical issues which must be addressed to effectively realize $\mu$-path scanning, which has only seen limited attention in the past.
% \end{enumerate}
\end{itemize}

@@

@@

\begin{figure}
  \includesvg[width=1\columnwidth]{figures/cs_raster_pixel_rows_6-5-2019.svg}
  \caption{Rows of pixels, as indicated by the red lines in
    Fig.~\ref{fig:resultsF1_images}. For clarity, not all images are
    included.}
  \label{fig:pixel_rows}
\end{figure}
For each raster scan and each CS scan, we computed the SSIM and PSNR
using the 1~Hz raster scan as the master and also computed the RDI
metric. These are plotted against total acquisition time in
Fig.~\ref{fig:time_vs_metrics}. Roughly speaking, image quality improves as
acquisition time increase. However, the metrics are quite noisy, leading to 
several inconsistencies. For example, it makes no sense that the 1~Hz, 25\% CS
scan has a \emph{worse} PSNR and SSIM than the 5~Hz, 25\% CS~scan;
nor does it make sense that the 8~Hz full raster scan has a worse SSIM and
PSNR than the 10~Hz full raster scan.

@@

\begin{figure}
  \includesvg[width=1\columnwidth]{figures/improvements.svg}
  \caption{Fraction of improvement in total scan time over a scan with
    512 lines vs scan rate for the two sets of CS scans (at 25\%
    and 12.5\%) and coarse raster scans at 128 and 64 lines.}
  \label{fig:improve_512}
\end{figure}

\section{Conclusions}\label{sec:conclusions}
In this paper, we introduced a new reconstruction method designed to reduce $\mu$-path
sampling artifacts. We then addressed several practical implementation issues.
Chief among these is the need to more aggressively mitigate coupling between the
horizontal plane and the $Z$-axis than is required for raster scanning. Here, we applied
a feedforward control scheme which attenuates the problematic modes by approximately 40 dB.

We then experimentally demonstrated $\mu$-path sub-sampling
and compared the resulting images to both full raster scans
and course raster scans paired with reconstruction.
The main conclusion of this comparison is that with $\mu$-paths, we
can acquire images faster yet with less specimen damage than is
possible while raster scanning a full image and that for slow scan rates,
$\mu$-path scanning achieves better images in less time than subline sampling.
On the other hand, for faster scan rates, reconstructing a coarse raster scan makes more sense.
This is notable, since a coarse raster scan with reconstruction is considerably
easier to implement, both on the hardware side as well as the post-processing side. 
It remains to be seen if this conclusion holds for a wide range of instruments.

%%%%%%%%% then the Bibliography, if any %%%%%%%%%
% \bibliographystyle{plain} % or "siam", or "alpha", etc.
\bibliographystyle{IEEEtran} % or "siam", or "alpha", etc.
\bibliography{braker_bibs,Andersson_bibs,Yufans_bibs} % Bib database in "refs.bib"
\vspace{-2em}
\begin{IEEEbiography}
  [{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{figures/braker_headshot.jpg}}]
  {Roger A. Braker} earned the B.A. degree in physics from the University of Oklahoma in 2005 and the M.S. degree in electrical engineering from the University of Colorado in 2017.
  Since 2014, he has been a graduate student in the Electrical, Computer, and Energy Engineering
  Department at the University of Colorado Boulder.
\end{IEEEbiography}
\begin{IEEEbiography}
  [{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{figures/yufan.JPG}}]
  {Yufan Luo}
received dual B.S. degrees in automation and in management science from Zhejiang University, Zhejiang, China, in 2011, an M.S. degree in industrial and operations engineering from the University of Michigan, Ann Arbor, MI, USA, in 2012, and a Ph.D. degree in systems engineering from Boston University in 2019. He is currently with Juniper Networks, Sunnyvale, CA.
\end{IEEEbiography}
\begin{IEEEbiography}
  [{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{figures/pao_headshot.jpg}}]
  {Lucy Y. Pao} 
  is the Palmer Endowed Chair Professor in the Electrical, Computer, and Energy Engineering Department at the University of Colorado Boulder. She earned B.S., M.S., and Ph.D. degrees in Electrical Engineering from Stanford University. Her research has primarily focused on engineering control systems, with applications ranging from atomic force microscopes to multi-megawatt wind energy systems. She is a Fellow of the IEEE and the International Federation of Automatic Control (IFAC). Selected recent awards include the 2012 IEEE Control Systems Magazine Outstanding Paper Award, the 2015 Society for Industrial and Applied Mathematics (SIAM) Journal on Control and Optimization Best Paper Prize, the 2017 Control Engineering Practice Award from the American Automatic Control Council, and the Scientific Award 2017 from the European Academy of Wind Energy. Selected recent and current professional society activities include being a Fellow of the Renewable and Sustainable Energy Institute (2009-present), 
  General Chair of the 2013 American Control Conference, member of the IEEE Control Systems Society (CSS) Board of Governors (2011-2013 and 2015), IEEE CSS Fellow Nominations Chair (2016-2018), and member of the IFAC Executive Board (2017-2020).    
\end{IEEEbiography}
\begin{IEEEbiography}
  [{\includegraphics[width=1 in,height=1.25in,clip,keepaspectratio]{figures/SeanAndersson_IEEE.jpg}}]
  {Sean B. Andersson}
received the B.S. degree in engineering and applied physics from Cornell University, Ithaca, NY, USA, in 1994, the M.S. degree in mechanical engineering from Stanford University, Stanford, CA, USA, in 1995, and the Ph.D. degree in electrical and computer engineering from the University of Maryland, College Park, MD, USA, in 2003. He has worked at AlliedSignal Aerospace and Aerovironment, Inc. and is currently an Associate Professor of mechanical engineering and of systems engineering with Boston University, Boston, MA, USA. His research interests include systems and control theory with applications in scanning probe microscopy, dynamics in molecular systems, and robotics. He received an NSF CAREER award in 2009, is a senior member of the IEEE, and was an associate editor for the IEEE Transactions on Automatic Control (2014-2018) and for the SIAM Journal on Control and Optimization (2013-2018).
\end{IEEEbiography}
